---
title: "frac_cover_tests"
output: html_document
date: "2024-02-01"
---

```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)

# visualization
library(mapview)
library(calecopal)

# significance tests and data exploration
library(tidyverse)
library(here)
library(kableExtra)
library(ggbeeswarm)
library(broom)
library(randomForest)
library(randomForestExplainer)
library(caret)
library(ggfortify)

# spatial analysis
library(sf)
library(tmap)
library(gstat)
library(stars)
library(terra)
library(SpatialKDE)
```

# Opening and exploring data

## Open soil and elevation data:

```{r}
#opening csv
soils <- read_csv(here("data", "shift_soil_data.csv")) %>% # read in soil csv
  mutate(date = lubridate::mdy(date), # change to date class
         week = lubridate::week(date)) %>% # create a week column
  select(soil_id:transect, meter_location, electro_cond_mS_per_cm, season, latitude, longitude, landcover)%>% #select certain column to work with
  mutate(paired_flight = lubridate::ymd(paired_flight)) %>% 
  drop_na() %>%  #drop any NA rows
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(value = 4326) %>% 
  st_transform(crs = 32611)


# Open spatial Data
soils_sf <- read_sf(here("data", "soil_w_elevation.shp")) %>% # read in sf file
  mutate(date = lubridate::ymd(date), #change date to date class
         week = lubridate::week(date)) %>% #create week
  select(electro_co, date, transect, meter_loca) %>% # select certain columns
  filter(date != "2022-09-12") %>% 
  mutate(electro_co = as.numeric(electro_co)) %>% #change to numeric class
  drop_na() #drop NAs

bbox <- read_sf(here("data", "bbox.shp")) #read in bounding box file

boundary <- read_sf(here("data", "boundary.shp")) %>% 
  st_transform(crs = 32611)

dem <- read_stars(here("data", "dem.tif")) %>% # read in dem 
  st_crop(y = st_bbox(bbox)) %>%  #crop to bounding box
  st_warp(cellsize = 4.8, crs = 32611)

mllw <- dem + 0.042

mllw_all <- c(mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20")), names = "date")

mllw_extract <- mllw %>% # call dem
  st_extract(soils$geometry) # extract elevation to point layer

```

## Join soil and elevation data:
```{r}
soils <- cbind(soils, mllw_extract$dem.tif) %>% #bind extracted values to the soil data
  rename("elevation" = "mllw_extract.dem.tif") %>% 
  st_transform(crs = 32611)
```

## Opening and Compiling NDVI:

```{r}
ndvi_1 <- read_stars(here("imagery", "NDVI", "2022_02_24_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_2 <- read_stars(here("imagery", "NDVI", "2022_02_28_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_3 <- read_stars(here("imagery", "NDVI", "2022_03_08_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_4 <- read_stars(here("imagery", "NDVI", "2022_03_16_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_5 <- read_stars(here("imagery", "NDVI", "2022_03_22_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_6 <- read_stars(here("imagery", "NDVI", "2022_04_05_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_7 <- read_stars(here("imagery", "NDVI", "2022_04_12_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_8 <- read_stars(here("imagery", "NDVI", "2022_04_20_ndvi.tif")) %>% 
  setNames("ndvi")
#ndvi_9 <- read_stars(here("imagery", "NDVI", "2022_04_29_ndvi.tif")) %>% 
#  setNames("ndvi")
#ndvi_10 <- read_stars(here("imagery", "NDVI", "2022_05_03_ndvi.tif")) %>% 
#  setNames("ndvi")
#ndvi_11 <- read_stars(here("imagery", "NDVI", "2022_05_11_ndvi.tif")) %>% 
#  setNames("ndvi")
#ndvi_12 <- read_stars(here("imagery", "NDVI", "2022_05_12_ndvi.tif")) %>% 
 # setNames("ndvi")
#ndvi_13 <- read_stars(here("imagery", "NDVI", "2022_05_17_ndvi.tif")) %>% 
#  setNames("ndvi")
#ndvi_14 <- read_stars(here("imagery", "NDVI", "2022_05_29_ndvi.tif")) %>% 
#  setNames("ndvi")


ndvi_all <- c(ndvi_1, ndvi_2, ndvi_3, ndvi_4, ndvi_5, ndvi_6, ndvi_7, ndvi_8, along = 3) %>% 
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20")), names = "date")
```

## Opening and Compiling VSSI:

```{r}
vssi_1 <- read_stars(here("imagery", "VSSI", "02_24_vssi.tif")) %>% 
  setNames("vssi")
vssi_2 <- read_stars(here("imagery", "VSSI", "02_28_vssi.tif")) %>% 
  setNames("vssi")
vssi_3 <- read_stars(here("imagery", "VSSI", "03_08_vssi.tif")) %>% 
  setNames("vssi")
vssi_4 <- read_stars(here("imagery", "VSSI", "03_16_vssi.tif")) %>% 
  setNames("vssi")
vssi_5 <- read_stars(here("imagery", "VSSI", "03_22_vssi.tif")) %>% 
  setNames("vssi")
vssi_6 <- read_stars(here("imagery", "VSSI", "04_05_vssi.tif")) %>% 
  setNames("vssi")
vssi_7 <- read_stars(here("imagery", "VSSI", "04_12_vssi.tif")) %>% 
  setNames("vssi")
vssi_8 <- read_stars(here("imagery", "VSSI", "04_20_vssi.tif")) %>% 
  setNames("vssi")
vssi_9 <- read_stars(here("imagery", "VSSI", "04_29_vssi.tif")) %>% 
  setNames("vssi")
vssi_10 <- read_stars(here("imagery", "VSSI", "05_03_vssi.tif")) %>% 
  setNames("vssi")
vssi_11 <- read_stars(here("imagery", "VSSI", "05_11_vssi.tif")) %>% 
  setNames("vssi")
#vssi_12 <- read_stars(here("imagery", "VSSI", "05_12_vssi.tif")) %>% 
 # setNames("vssi")
vssi_13 <- read_stars(here("imagery", "VSSI", "05_17_vssi.tif")) %>% 
  setNames("vssi")
vssi_14 <- read_stars(here("imagery", "VSSI", "05_29_vssi")) %>% 
  setNames("vssi")

vssi_all <- c(vssi_1, vssi_2, vssi_3, vssi_4, vssi_5, vssi_6, vssi_7, vssi_8, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20")), names = "date")
```

## Opening and Compiling mARI:

```{r}
mari_1 <- read_stars(here("imagery", "mARI", "02_24_mari_avg.tif")) %>% 
  setNames("mari")
mari_2 <- read_stars(here("imagery", "mARI", "02_28_mari_avg.tif")) %>% 
  setNames("mari")
mari_3 <- read_stars(here("imagery", "mARI", "03_08_mari_avg.tif")) %>% 
  setNames("mari")
mari_4 <- read_stars(here("imagery", "mARI", "03_16_mari_avg.tif")) %>% 
  setNames("mari")
mari_5 <- read_stars(here("imagery", "mARI", "03_22_mari_avg.tif")) %>% 
  setNames("mari")
mari_6 <- read_stars(here("imagery", "mARI", "04_05_mari_avg.tif")) %>% 
  setNames("mari")
mari_7 <- read_stars(here("imagery", "mARI", "04_12_mari_avg.tif")) %>% 
  setNames("mari")
mari_8 <- read_stars(here("imagery", "mARI", "04_20_mari_avg.tif")) %>% 
  setNames("mari")
mari_9 <- read_stars(here("imagery", "mARI", "04_29_mari_avg.tif")) %>% 
  setNames("mari")
mari_10 <- read_stars(here("imagery", "mARI", "05_03_mari_avg.tif")) %>% 
  setNames("mari")
mari_11 <- read_stars(here("imagery", "mARI", "05_11_mari_avg.tif")) %>% 
  setNames("mari")
#mari_12 <- read_stars(here("imagery", "mARI", "05_12_mari_avg.tif")) %>% 
 # setNames("mari")
mari_13 <- read_stars(here("imagery", "mARI", "05_17_mari_avg.tif")) %>% 
  setNames("mari")
mari_14 <- read_stars(here("imagery", "mARI", "05_29_mari_avg.tif")) %>% 
  setNames("mari")

mari_all <- c(mari_1, mari_2, mari_3, mari_4, mari_5, mari_6, mari_7, mari_8, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20")), names = "date")
```

## Opening and Compiling CRSI:

```{r}
crsi_1 <- read_stars(here("imagery", "CRSI", "02_24_crsi.tif")) %>% 
  setNames("crsi")
crsi_2 <- read_stars(here("imagery", "CRSI", "02_28_crsi.tif")) %>% 
  setNames("crsi")
crsi_3 <- read_stars(here("imagery", "CRSI", "03_08_crsi.tif")) %>% 
  setNames("crsi")
crsi_4 <- read_stars(here("imagery", "CRSI", "03_16_crsi.tif")) %>% 
  setNames("crsi")
crsi_5 <- read_stars(here("imagery", "CRSI", "03_22_crsi.tif")) %>% 
  setNames("crsi")
crsi_6 <- read_stars(here("imagery", "CRSI", "04_05_crsi.tif")) %>% 
  setNames("crsi")
crsi_7 <- read_stars(here("imagery", "CRSI", "04_12_crsi.tif")) %>% 
  setNames("crsi")
crsi_8 <- read_stars(here("imagery", "CRSI", "04_20_crsi.tif")) %>% 
  setNames("crsi")
crsi_9 <- read_stars(here("imagery", "CRSI", "04_29_crsi.tif")) %>% 
  setNames("crsi")
crsi_10 <- read_stars(here("imagery", "CRSI", "05_03_crsi.tif")) %>% 
  setNames("crsi")
crsi_11 <- read_stars(here("imagery", "CRSI", "05_11_crsi.tif")) %>% 
  setNames("crsi")
#crsi_12 <- read_stars(here("imagery", "CRSI", "05_12_crsi.tif")) %>% 
 # setNames("crsi")
crsi_13 <- read_stars(here("imagery", "CRSI", "05_17_crsi.tif")) %>% 
  setNames("crsi")
crsi_14 <- read_stars(here("imagery", "CRSI", "05_29_crsi.tif")) %>% 
  setNames("crsi")

crsi_all <- c(crsi_1, crsi_2, crsi_3, crsi_4, crsi_5, crsi_6, crsi_7, crsi_8, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20")), names = "date")
```

# G/N/S Fractions

```{r}
#frac_7 <- read_stars(here("imagery", "subset_images", "fractions_test","rfl_2022_04_20.tif_SMA_20240131T15h30M20S")) %>% 
#  filter(band %in% c("NPV_fraction", "GREEN VEG_fraction", "BARE SOIL_fraction")) %>% 
#  split(3) %>% 
#  setNames(c("NPV_fraction", "GREEN_VEG_fraction", "BARE_SOIL_fraction"))

frac_8 <- read_stars(here("imagery", "subset_images", "fractions_test","rfl_2022_04_20.tif_SMA_20240131T18h40M07S")) %>% 
  filter(band %in% c("NPV_fraction", "GREEN VEG_fraction", "BARE SOIL_fraction")) %>% 
  split(3) %>% 
  setNames(c("NPV_fraction", "GREEN_VEG_fraction", "BARE_SOIL_fraction"))

#frac_all <- c(frac_7, frac_8, along = 3) %>%
#  st_set_dimensions(3, values = lubridate::ymd(c("2022_04_12", "2022_04_20")), names = "date") 

npv_frac_8 <- frac_8 %>% 
  select("NPV_fraction")

gv_frac_8 <- frac_8 %>% 
  select("GREEN_VEG_fraction")

soil_frac_8 <- frac_8 %>% 
  select("BARE_SOIL_fraction")


#npv_frac_all <- frac_all %>% 
#  select("NPV_fraction")

#gv_frac_all <- frac_all %>% 
#  select("GREEN_VEG_fraction")

#soil_frac_all <- frac_all %>% 
#  select("BARE_SOIL_fraction")

```

# TESTING FRACTIONAL COVER INCLUSION

## EXTRACT FRACTIONAL COVER

```{r}
soils <- soils %>% 
  filter(paired_flight == "2022-04-20")

rast_all <- c(ndvi_8, mari_8, crsi_8, vssi_8, npv_frac_8, gv_frac_8, soil_frac_8)

rast_extract <- rast_all %>% 
  st_extract(soils) %>% 
  st_drop_geometry()

soils_extract <- soils %>% 
  bind_cols(rast_extract)

```

## FIT MODEL

```{r}
set.seed(1234)

fit_all_frac <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + ndvi + crsi + GREEN_VEG_fraction + BARE_SOIL_fraction + NPV_fraction,
                      data = soils_extract, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + crsi + GREEN_VEG_fraction + BARE_SOIL_fraction + NPV_fraction,
                      data = soils_extract) %>% 
  car::vif()

```

## EVALUATE MODEL

```{r}
# using caret
fit_all_frac_imp <- as.data.frame(fit_all_frac$finalModel$importance)
fit_all_frac_imp_scaled <- predict(preProcess(fit_all_frac_imp, method = c("range")), fit_all_frac_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_frac_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_frac_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_frac_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(fit_all_frac$finalModel$y, fit_all_frac$finalModel$predicted)
lm(fit_all_frac$finalModel$y ~ fit_all_frac$finalModel$predicted)
```


## PLOT RELATIONSHIP

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation, Fractional Cover, and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

all_corr <- cor(y= fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted)
all_corr
```




### COMPARISON

## FIT MODEL

```{r}
set.seed(1234)

fit_all <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + ndvi + crsi,
                      data = soils_extract, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

## EVALUATE MODEL

```{r}
# using caret
fit_all_imp <- as.data.frame(fit_all$finalModel$importance)
fit_all_imp_scaled <- predict(preProcess(fit_all_imp, method = c("range")), fit_all_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(fit_all$finalModel$y, fit_all$finalModel$predicted)
lm(fit_all$finalModel$y ~ fit_all$finalModel$predicted)
```


## PLOT RELATIONSHIP

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_all$finalModel$y, x = fit_all$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all$finalModel$y, x = fit_all$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation, Fractional Cover, and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

all_corr <- cor(y= fit_all$finalModel$y, x = fit_all$finalModel$predicted)
all_corr
```


