---
title: "Fractional Cover Tests"
output: html_document
date: "2023-11-29"
---
```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)

# visualization
library(mapview)
library(calecopal)

# significance tests and data exploration
library(tidyverse)
library(here)
library(kableExtra)
library(ggbeeswarm)
library(broom)
library(randomForest)
library(randomForestExplainer)
library(caret)
library(ggfortify)

# spatial analysis
library(sf)
library(tmap)
library(gstat)
library(stars)
library(terra)
library(SpatialKDE)
```

# Opening and exploring data

## Open soil and elevation data:

```{r}
#opening csv
soils <- read_csv(here("data", "shift_soil_data.csv")) %>% # read in soil csv
  mutate(date = lubridate::mdy(date), # change to date class
         week = lubridate::week(date)) %>% # create a week column
  select(soil_id:transect, meter_location, electro_cond_mS_per_cm, season, latitude, longitude, landcover)%>% #select certain column to work with
  mutate(paired_flight = lubridate::ymd(paired_flight)) %>% 
  drop_na() %>%  #drop any NA rows
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(value = 4326) %>% 
  st_transform(crs = 32611)


# Open spatial Data
soils_sf <- read_sf(here("data", "soil_w_elevation.shp")) %>% # read in sf file
  mutate(date = lubridate::ymd(date), #change date to date class
         week = lubridate::week(date)) %>% #create week
  select(electro_co, date, transect, meter_loca) %>% # select certain columns
  filter(date != "2022-09-12") %>% 
  mutate(electro_co = as.numeric(electro_co)) %>% #change to numeric class
  drop_na() #drop NAs

bbox <- read_sf(here("data", "bbox.shp")) #read in bounding box file

boundary <- read_sf(here("data", "marsh_boundary.shp")) %>% 
  st_transform(crs = 32611)

dem <- read_stars(here("data", "dem.tif")) %>% # read in dem 
  st_crop(y = st_bbox(bbox)) %>%  #crop to bounding box
  st_warp(cellsize = 4.8, crs = 32611)

mllw <- dem + 0.042

mllw_all <- c(mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")


mllw_extract <- mllw %>% # call dem
  st_extract(soils$geometry) # extract elevation to point layer

```


## Join soil and elevation data:
```{r}
soils <- cbind(soils, mllw_extract$dem.tif) %>% #bind extracted values to the soil data
  rename("elevation" = "mllw_extract.dem.tif") %>% 
  st_transform(crs = 32611)
```

# Open Raster Files:


```{r}
savi_1 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_24.tif")) %>% 
  st_set_dimensions(3, values = seq(1:425)) %>% 
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>% 
  select(savi)

savi_2 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_28.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_3 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_08.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_4 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_16.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_5 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_22.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_6 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_05.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_7 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_12.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_8 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_20.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_9 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_10 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_03.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_11 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_11.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_13 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_17.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_14 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  split(3) %>%
  mutate(savi = ((((X85+X86)/2)-((X59+X60)/2))/(((X85+X86)/2)+((X59+X60)/2)+0.5)) * 1.5) %>%
  select(savi)

savi_all <- c(savi_1, savi_2, savi_3, savi_4, savi_5, savi_6, savi_7, savi_8, savi_9, savi_10, savi_11, savi_13, savi_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```


## Opening and Compiling NDVI:
```{r}
ndvi_1 <- read_stars(here("imagery", "NDVI", "2022_02_24_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_2 <- read_stars(here("imagery", "NDVI", "2022_02_28_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_3 <- read_stars(here("imagery", "NDVI", "2022_03_08_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_4 <- read_stars(here("imagery", "NDVI", "2022_03_16_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_5 <- read_stars(here("imagery", "NDVI", "2022_03_22_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_6 <- read_stars(here("imagery", "NDVI", "2022_04_05_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_7 <- read_stars(here("imagery", "NDVI", "2022_04_12_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_8 <- read_stars(here("imagery", "NDVI", "2022_04_20_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_9 <- read_stars(here("imagery", "NDVI", "2022_04_29_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_10 <- read_stars(here("imagery", "NDVI", "2022_05_03_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_11 <- read_stars(here("imagery", "NDVI", "2022_05_11_ndvi.tif")) %>% 
  setNames("ndvi")
#ndvi_12 <- read_stars(here("imagery", "NDVI", "2022_05_12_ndvi.tif")) %>% 
 # setNames("ndvi")
ndvi_13 <- read_stars(here("imagery", "NDVI", "2022_05_17_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_14 <- read_stars(here("imagery", "NDVI", "2022_05_29_ndvi.tif")) %>% 
  setNames("ndvi")


ndvi_all <- c(ndvi_1, ndvi_2, ndvi_3, ndvi_4, ndvi_5, ndvi_6, ndvi_7, ndvi_8, ndvi_9, ndvi_10, ndvi_11, ndvi_13, ndvi_14, along = 3) %>% 
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling VSSI:

```{r}
vssi_1 <- read_stars(here("imagery", "VSSI", "02_24_vssi.tif")) %>% 
  setNames("vssi")
vssi_2 <- read_stars(here("imagery", "VSSI", "02_28_vssi.tif")) %>% 
  setNames("vssi")
vssi_3 <- read_stars(here("imagery", "VSSI", "03_08_vssi.tif")) %>% 
  setNames("vssi")
vssi_4 <- read_stars(here("imagery", "VSSI", "03_16_vssi.tif")) %>% 
  setNames("vssi")
vssi_5 <- read_stars(here("imagery", "VSSI", "03_22_vssi.tif")) %>% 
  setNames("vssi")
vssi_6 <- read_stars(here("imagery", "VSSI", "04_05_vssi.tif")) %>% 
  setNames("vssi")
vssi_7 <- read_stars(here("imagery", "VSSI", "04_12_vssi.tif")) %>% 
  setNames("vssi")
vssi_8 <- read_stars(here("imagery", "VSSI", "04_20_vssi.tif")) %>% 
  setNames("vssi")
vssi_9 <- read_stars(here("imagery", "VSSI", "04_29_vssi.tif")) %>% 
  setNames("vssi")
vssi_10 <- read_stars(here("imagery", "VSSI", "05_03_vssi.tif")) %>% 
  setNames("vssi")
vssi_11 <- read_stars(here("imagery", "VSSI", "05_11_vssi.tif")) %>% 
  setNames("vssi")
#vssi_12 <- read_stars(here("imagery", "VSSI", "05_12_vssi.tif")) %>% 
 # setNames("vssi")
vssi_13 <- read_stars(here("imagery", "VSSI", "05_17_vssi.tif")) %>% 
  setNames("vssi")
vssi_14 <- read_stars(here("imagery", "VSSI", "05_29_vssi")) %>% 
  setNames("vssi")

vssi_all <- c(vssi_1, vssi_2, vssi_3, vssi_4, vssi_5, vssi_6, vssi_7, vssi_8, vssi_9, vssi_10, vssi_11, vssi_13, vssi_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling mARI:

```{r}
mari_1 <- read_stars(here("imagery", "mARI", "02_24_mari_avg.tif")) %>% 
  setNames("mari")
mari_2 <- read_stars(here("imagery", "mARI", "02_28_mari_avg.tif")) %>% 
  setNames("mari")
mari_3 <- read_stars(here("imagery", "mARI", "03_08_mari_avg.tif")) %>% 
  setNames("mari")
mari_4 <- read_stars(here("imagery", "mARI", "03_16_mari_avg.tif")) %>% 
  setNames("mari")
mari_5 <- read_stars(here("imagery", "mARI", "03_22_mari_avg.tif")) %>% 
  setNames("mari")
mari_6 <- read_stars(here("imagery", "mARI", "04_05_mari_avg.tif")) %>% 
  setNames("mari")
mari_7 <- read_stars(here("imagery", "mARI", "04_12_mari_avg.tif")) %>% 
  setNames("mari")
mari_8 <- read_stars(here("imagery", "mARI", "04_20_mari_avg.tif")) %>% 
  setNames("mari")
mari_9 <- read_stars(here("imagery", "mARI", "04_29_mari_avg.tif")) %>% 
  setNames("mari")
mari_10 <- read_stars(here("imagery", "mARI", "05_03_mari_avg.tif")) %>% 
  setNames("mari")
mari_11 <- read_stars(here("imagery", "mARI", "05_11_mari_avg.tif")) %>% 
  setNames("mari")
#mari_12 <- read_stars(here("imagery", "mARI", "05_12_mari_avg.tif")) %>% 
 # setNames("mari")
mari_13 <- read_stars(here("imagery", "mARI", "05_17_mari_avg.tif")) %>% 
  setNames("mari")
mari_14 <- read_stars(here("imagery", "mARI", "05_29_mari_avg.tif")) %>% 
  setNames("mari")

mari_all <- c(mari_1, mari_2, mari_3, mari_4, mari_5, mari_6, mari_7, mari_8, mari_9, mari_10, mari_11, mari_13, mari_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling CRSI:

```{r}
crsi_1 <- read_stars(here("imagery", "CRSI", "02_24_crsi.tif")) %>% 
  setNames("crsi")
crsi_2 <- read_stars(here("imagery", "CRSI", "02_28_crsi.tif")) %>% 
  setNames("crsi")
crsi_3 <- read_stars(here("imagery", "CRSI", "03_08_crsi.tif")) %>% 
  setNames("crsi")
crsi_4 <- read_stars(here("imagery", "CRSI", "03_16_crsi.tif")) %>% 
  setNames("crsi")
crsi_5 <- read_stars(here("imagery", "CRSI", "03_22_crsi.tif")) %>% 
  setNames("crsi")
crsi_6 <- read_stars(here("imagery", "CRSI", "04_05_crsi.tif")) %>% 
  setNames("crsi")
crsi_7 <- read_stars(here("imagery", "CRSI", "04_12_crsi.tif")) %>% 
  setNames("crsi")
crsi_8 <- read_stars(here("imagery", "CRSI", "04_20_crsi.tif")) %>% 
  setNames("crsi")
crsi_9 <- read_stars(here("imagery", "CRSI", "04_29_crsi.tif")) %>% 
  setNames("crsi")
crsi_10 <- read_stars(here("imagery", "CRSI", "05_03_crsi.tif")) %>% 
  setNames("crsi")
crsi_11 <- read_stars(here("imagery", "CRSI", "05_11_crsi.tif")) %>% 
  setNames("crsi")
#crsi_12 <- read_stars(here("imagery", "CRSI", "05_12_crsi.tif")) %>% 
 # setNames("crsi")
crsi_13 <- read_stars(here("imagery", "CRSI", "05_17_crsi.tif")) %>% 
  setNames("crsi")
crsi_14 <- read_stars(here("imagery", "CRSI", "05_29_crsi.tif")) %>% 
  setNames("crsi")

crsi_all <- c(crsi_1, crsi_2, crsi_3, crsi_4, crsi_5, crsi_6, crsi_7, crsi_8, crsi_9, crsi_10, crsi_11, crsi_13, crsi_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```


## Calling Fractional Cover Data

### All Fractions

```{r}
frac_1 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_02_24_shadeNorm")) %>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_2 <-read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_02_28_shadeNorm")) %>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_3 <-read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_03_08_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_4 <-read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_03_16_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_5 <-read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_03_22_shadeNorm")) %>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_6 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_04_05_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_7 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_04_12_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_8 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_04_20_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_9 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_04_29_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_10 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_05_03_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_11 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_05_11_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_13 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_05_17_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_14 <- read_stars(here("imagery", "fractional_cover", "shade_norm", "rfl_2022_05_29_shadeNorm"))%>% 
  filter(band %in% c("ShadeNorm_ NPV_fraction", "ShadeNorm_ GREEN VEG_fraction", "ShadeNorm_ BARE SOIL_fraction")) %>%
  split(3) %>% 
  setNames(c("Soil_fraction", "GV_fraction", "NPV_fraction" ))

frac_all <- c(frac_1, frac_2, frac_3, frac_4, frac_5, frac_6, frac_7, frac_8, frac_9, frac_10, frac_11, frac_13, frac_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

### NPV

```{r}
npv_frac_all <- frac_all %>% 
  select("NPV_fraction")
```

### GV

```{r}
gv_frac_all <- frac_all %>% 
  select("GV_fraction")
```

### Bare Soil

```{r}
soil_frac_all <- frac_all %>% 
  select("Soil_fraction")
```

## Adjusting elevation data

```{r}
mllw_all <- mllw_all %>% 
  st_warp(dest = crsi_all) %>% 
  setNames("elevation")

savi_all <- savi_all %>% 
  st_warp(dest = crsi_all) %>% 
  st_as_stars()

```

# TESTING FRACTIONAL COVER INCLUSION

## EXTRACT FRACTIONAL COVER

```{r}

soils_subset <- soils %>% 
  filter(paired_flight %in% lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")))

soils_extract_2 <- data.frame()

for (i in unique(lubridate::week(soils_subset$paired_flight))){
  ndvi_week <- ndvi_all %>% 
    filter(lubridate::week(date) == i)
  
  mari_week <- mari_all %>% 
    filter(lubridate::week(date) == i)

  vssi_week <- vssi_all %>% 
    filter(lubridate::week(date) == i)
  
  crsi_week <- crsi_all %>% 
    filter(lubridate::week(date) == i)
  
  frac_week <- frac_all %>% 
    filter(lubridate::week(date) == i) %>% 
    st_as_sf()
  
  savi_week <- savi_all %>% 
    filter(lubridate::week(date) == i)
  
  soil_week <- soils %>% 
    mutate(week = lubridate::week(paired_flight)) %>% 
    filter(week == i)
  
  ndvi_extract <- ndvi_week %>% 
    st_extract(soil_week)
  
  mari_extract <- mari_week %>% 
    st_extract(soil_week)
  
  vssi_extract <- vssi_week %>% 
    st_extract(soil_week)
  
  crsi_extract <- crsi_week %>% 
    st_extract(soil_week)
  
  frac_extract <- soil_week %>% 
    st_join(frac_week) %>% 
    select("Soil_fraction":"NPV_fraction") %>% 
    st_drop_geometry()
  
  savi_extract <- savi_week %>% 
   st_extract(soil_week)
  
  soils_week <- bind_cols(soil_week, ndvi_extract$ndvi, mari_extract$mari, vssi_extract$vssi, crsi_extract$crsi, savi_extract$savi, frac_extract) %>% 
    rename("ndvi" = "...12",
           "mari" = "...13",
           "vssi" = "...14",
           "crsi" = "...15",
           "savi" = "...16")
    
  soils_extract_2 <- rbind(soils_extract_2, soils_week)
}

#write.csv(soils_extract, file = here("data", "extracted_soils.csv"))
```

## FIT MODEL

```{r}
set.seed(1234)

fit_all_savi <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + savi,
                      data = soils_extract_2, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

## EVALUATE MODEL

```{r}
# using caret
fit_all_savi_imp <- as.data.frame(fit_all_savi$finalModel$importance)
fit_all_savi_imp_scaled <- predict(preProcess(fit_all_savi_imp, method = c("range")), fit_all_savi_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_savi_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Increase in Node Purity")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))


ggplot(fit_all_savi_imp) +
  geom_point(aes(x = fit_all_savi_imp$`%IncMSE`, y = row.names(fit_all_savi_imp)), color ="#3A5D3D")+ 
  geom_segment(aes(y = row.names(fit_all_savi_imp), x = 0, yend = row.names(fit_all_savi_imp), xend = fit_all_savi_imp$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (%)")+
  ggtitle("% Increase in MSE")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

## PLOT RELATIONSHIP

```{r}
ggplot() +#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_point(aes(y = fit_all_savi$finalModel$y, x = fit_all_savi$finalModel$predicted), color = "#E69512") +#data to make points from
  geom_smooth(method = "lm", aes(y = fit_all_savi$finalModel$y, x = fit_all_savi$finalModel$predicted), color = "#3A5D3D") +#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14)) +
  labs(y = "Actual Electrical Conductivity (mS/cm)", #labels
       x = "Predicted Electrical Conductivity (mS/cm)") +
  ggtitle("Elevation and Indices") +#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5), #plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```


```{r}
set.seed(1234)

fit_all_ele <- train(electro_cond_mS_per_cm ~ elevation,
                      data = soils_extract_2, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```


## PLOT RELATIONSHIP

```{r}
ggplot() +#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted") +
  geom_point(aes(y = fit_all_ele$finalModel$y, x = fit_all_ele$finalModel$predicted), color = "#E69512") +#data to make points from
  geom_smooth(method = "lm", aes(y = fit_all_ele$finalModel$y, x = fit_all_ele$finalModel$predicted), color = "#3A5D3D") +#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14)) +
  labs(y = "Actual Electrical Conductivity (mS/cm)", #labels
       x = "Predicted Electrical Conductivity (mS/cm)") +
  ggtitle("Elevation") +#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5), #plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

### Correlation and predicted fit metrics

```{r}
all_corr <- cor(y= fit_all_ele$finalModel$y, x = fit_all_ele$finalModel$predicted)
all_corr

lm(fit_all_ele$finalModel$y ~ fit_all_ele$finalModel$predicted)
max(fit_all_ele$finalModel$predicted)
min(fit_all_ele$finalModel$predicted)
```

# Fitting model to all data

```{r}
raster_all <- c(crsi_all, ndvi_all, vssi_all, mari_all, mllw_all, savi_all)

all_pred_savi <- predict(raster_all, fit_all_savi, drop_dimensions = F, na.action = na.omit)

#write_stars(all_pred_savi, dsn = here("data", "prediction_savi.tif"))
```

# Making a Map

```{r}
all_pred_crop <- all_pred_savi %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE) %>% 
  filter(date == "2022-02-24")


ggplot()+
  geom_stars(data = all_pred_crop, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "kelp1", n = 4, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Latitude",#labels
       x = "Longitude",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```




##################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################

## NON COLLINEAR AND NO FRACTIONS

```{r}
set.seed(1234)

fit_all_nc <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + ndvi,
                      data = soils_extract_2, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)
```

```{r}
# using caret
fit_all_nc_imp <- as.data.frame(fit_all_nc$finalModel$importance)
fit_all_nc_imp_scaled <- predict(preProcess(fit_all_nc_imp, method = c("range")), fit_all_nc_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_nc_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_nc_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_nc_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

```


```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation and Spectral Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```


```{r}
all_corr <- cor(y= fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted)
all_corr

lm(fit_all_nc$finalModel$y ~ fit_all_nc$finalModel$predicted)
max(fit_all_nc$finalModel$predicted)
min(fit_all_nc$finalModel$predicted)
```

```{r}
all_pred_nc <- predict(raster_all, fit_all_nc, drop_dimensions = F, na.action = na.omit)

plot(all_pred_nc)
```



```{r}

all_pred_crop <- all_pred_nc  %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE) %>% 
  filter(date == "2022-02-24")

#write_stars(all_pred_crop, dsn = here("data", "prediction_ndvi.tif"))

ggplot()+
  geom_stars(data = all_pred_crop, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "kelp1", n = 4, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Latitude",#labels
       x = "Longitude",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```



# Comparing

```{r}
t.test(x = fit_all_nc$finalModel$predicted, y = fit_all_savi$finalModel$predicted)

effsize::cohen.d(fit_all_nc$finalModel$predicted, fit_all_savi$finalModel$predicted)
```







##################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################

# TESTING COLLINEARITY AND ALTERATIVE MODEL FITS

## Collinearity

### All variables

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + savi, data = soils_extract_2)

car::vif(model)

```

### Remove NDVI

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + savi + GV_fraction, data = soils_extract_2)

car::vif(model)

```

No real change in GV fraction, Soil is still high, perhaps it was soil fractions in actuality

### Remove Soil fractions, keep ndvi

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + ndvi + NPV_fraction + GV_fraction, data = soils_extract_2)

car::vif(model)

```
GV dropped but MARI and NDVI increased, lets remove GV and/or ndvi

```{r}
# remove ndvi
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + GV_fraction, data = soils_extract_2)

car::vif(model)
```

```{r}
# Add Soil remove GV
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi + Soil_fraction, data = soils_extract_2)

car::vif(model)
```
removing ndvi and leaving GV fractions leaves acceptable collinearity



## FIT MODEL

### SOIL FRACTIONS ONLY

```{r}
set.seed(1234)

fit_all_frac <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + Soil_fraction,
                      data = soils_extract_2, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

#### EVALUATE MODEL

```{r}
# using caret
fit_all_frac_imp <- as.data.frame(fit_all_frac$finalModel$importance)
fit_all_frac_imp_scaled <- predict(preProcess(fit_all_frac_imp, method = c("range")), fit_all_frac_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_frac_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_frac_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_frac_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

#### PLOT RELATIONSHIP

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation, Fractional Cover, and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```


#### Correlation and predicted fit metrics

```{r}
all_corr <- cor(y= fit_all_frac$finalModel$y, x = fit_all_frac$finalModel$predicted)
all_corr

lm(fit_all_frac$finalModel$y ~ fit_all_frac$finalModel$predicted)
max(fit_all_frac$finalModel$predicted)
min(fit_all_frac$finalModel$predicted)
```

##### Fitting model to all data

```{r}
raster_all <- c(crsi_all, ndvi_all, vssi_all, mari_all, mllw_all, gv_frac_all, soil_frac_all, npv_frac_all)

all_pred_frac <- predict(raster_all, fit_all_frac, drop_dimensions = F, na.action = na.omit)

```

#### Making a Map

```{r}
all_pred_crop <- all_pred_frac %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE) %>% 
  filter(date == "2022-02-24")



#write_stars(all_pred_crop, dsn = here("data", "predictions_nc_feb.tif"))


ggplot()+
  geom_stars(data = all_pred_crop, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "kelp1", n = 4, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Latitude",#labels
       x = "Longitude",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```


### GV FRACTIONS ONLY

```{r}
set.seed(1234)

fit_all_gv <- train(electro_cond_mS_per_cm ~ elevation + mari + vssi + GV_fraction,
                      data = soils_extract_2, 
                      method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

## EVALUATE MODEL

```{r}
# using caret
fit_all_gv_imp <- as.data.frame(fit_all_gv$finalModel$importance)
fit_all_gv_imp_scaled <- predict(preProcess(fit_all_gv_imp, method = c("range")), fit_all_gv_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_gv_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_gv_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_gv_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

## PLOT RELATIONSHIP

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_all_gv$finalModel$y, x = fit_all_gv$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all_gv$finalModel$y, x = fit_all_gv$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation, GV Fractional Cover, and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```


### Correlation and predicted fit metrics

```{r}
all_corr <- cor(y= fit_all_gv$finalModel$y, x = fit_all_gv$finalModel$predicted)
all_corr

lm(fit_all_gv$finalModel$y ~ fit_all_gv$finalModel$predicted)
max(fit_all_gv$finalModel$predicted)
min(fit_all_gv$finalModel$predicted)
```

# Fitting model to all data

```{r}
all_pred_gv <- predict(raster_all, fit_all_gv, drop_dimensions = F, na.action = na.omit)

plot(all_pred_gv)

```

# Making a Map

```{r}
all_pred_crop <- all_pred_frac %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE) %>% 
  filter(date == "2022-02-24")



#write_stars(all_pred_crop, dsn = here("data", "predictions_nc_feb.tif"))


ggplot()+
  geom_stars(data = all_pred_crop, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "kelp1", n = 4, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Latitude",#labels
       x = "Longitude",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```


















