---
title: "Soil Data Exploration"
author: "German Silva"
date: "3/4/2022"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)

# visualization
library(mapview)
library(calecopal)

# significance tests and data exploration
library(tidyverse)
library(here)
library(kableExtra)
library(ggbeeswarm)
library(broom)
library(randomForest)
library(randomForestExplainer)
library(caret)
library(ggfortify)

# spatial analysis
library(sf)
library(tmap)
library(gstat)
library(stars)
library(terra)
library(SpatialKDE)

set.seed(1234)
```

# Opening and exploring data

## Open soil and elevation data:

```{r}
#opening csv
soils <- read_csv(here("data", "shift_soil_data.csv")) %>% # read in soil csv
  mutate(date = lubridate::mdy(date), # change to date class
         week = lubridate::week(date)) %>% # create a week column
  select(soil_id:transect, meter_location, electro_cond_mS_per_cm, season, latitude, longitude, landcover)%>% #select certain column to work with
  mutate(paired_flight = lubridate::ymd(paired_flight)) %>% 
  drop_na() %>%  #drop any NA rows
  st_as_sf(coords = c("longitude", "latitude")) %>% 
  st_set_crs(value = 4326) %>% 
  st_transform(crs = 32611)


# Open spatial Data
soils_sf <- read_sf(here("data", "soil_w_elevation.shp")) %>% # read in sf file
  mutate(date = lubridate::ymd(date), #change date to date class
         week = lubridate::week(date)) %>% #create week
  select(electro_co, date, transect, meter_loca) %>% # select certain columns
  filter(date != "2022-09-12") %>% 
  mutate(electro_co = as.numeric(electro_co)) %>% #change to numeric class
  drop_na() #drop NAs

bbox <- read_sf(here("data", "bbox.shp")) #read in bounding box file

boundary <- read_sf(here("data", "boundary.shp")) %>% 
  st_transform(crs = 32611)

dem <- read_stars(here("data", "dem.tif")) %>% # read in dem 
  st_crop(y = st_bbox(bbox)) %>%  #crop to bounding box
  st_warp(cellsize = 4.8, crs = 32611)

mllw <- dem + 0.042

mllw_all <- c(mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, mllw, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")

mllw_extract <- mllw %>% # call dem
  st_extract(soils$geometry) # extract elevation to point layer

```


## Join soil and elevation data:
```{r}
soils <- cbind(soils, mllw_extract$dem.tif) %>% #bind extracted values to the soil data
  rename("elevation" = "mllw_extract.dem.tif") %>% 
  st_transform(crs = 32611)
```

## Descriptive statistics:
```{r}
#make descriptive stat table
soils %>% # to soils
  group_by(transect) %>%# group by transect  
  summarize(min = min(electro_cond_mS_per_cm),#calculate mins
            max = max(electro_cond_mS_per_cm),#calculate max
            mean = mean(electro_cond_mS_per_cm),#calculate mean
            median = median(electro_cond_mS_per_cm),
            dev = sd(electro_cond_mS_per_cm),#obtain standard dev
            variance = var(electro_cond_mS_per_cm)) %>% # calculate variance
  kableExtra::kable() %>% # create a table format
  kableExtra::kable_classic(lightable_options = "striped")#theme the table

# make descriptive stat dataframe
soil_stats <- soils %>% #call soils
  group_by(transect) %>% #group by transects
  summarize(mean = mean(electro_cond_mS_per_cm),# create dataframe with these stats
            max = max(electro_cond_mS_per_cm),
            min = min(electro_cond_mS_per_cm),
            range = max - min)
```

## Plot salinity v elevation
```{r}
ggplot(soils, aes(x = elevation, y = electro_cond_mS_per_cm, color = transect)) +#ggplot of soils/dem data
  geom_point()+#geometry of the plot
  scale_color_manual(values = calecopal::cal_palette(name = "superbloom3", n =6, type = "discrete"))+# colors to use
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")+
  theme(legend.position = "none")+#no legend theme
  guides(color = guide_legend(title = "Transect ID"))+#changing legend title
  ggtitle("Soil Electrical Conductivity Across Elevations")+# title of plot
  theme(plot.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),# plot theming
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 14),
        axis.title = element_text(color = "#5b4f41", size = 16),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"))
```

## Salinity vs. elevation, facet wrapped
```{r}
strip_labels <- c("COPR_1_EW" = "COPR 1 EW",#create better formated labels for facet wrap strips
                  "COPR_2_EW" = "COPR 2 EW",
                  "COPR_2_NS" = "COPR 2 NS",
                  "NCOS_1_NS" = "NCOS 1 NS",
                  "NCOS_2_EW" = "NCOS 2 EW",
                  "NCOS_2_NS" = "NCOS 2 NS")

ggplot(soils, aes(x = elevation, y = electro_cond_mS_per_cm, color = transect)) +#start ggplot
  geom_point()+# point geometry
  scale_color_manual(values = calecopal::cal_palette(name = "superbloom3", n =6, type = "discrete"))+#colors
  facet_wrap(~transect, labeller = as_labeller(strip_labels))+#subplots based on transect
  theme(legend.position = "none")+# legend theme
  labs(x = "Elevation (m)",#labels
       y = "Electrical conductivity (mS/cm)")+
  guides(color = guide_legend(title = "Transect ID"))+#legend title
  ggtitle("Soil Electrical Conductivity Across Elevations")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot theme
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"))
```

# Open Raster Files:

## Opening and Compiling NDVI:

```{r}
ndvi_1 <- read_stars(here("imagery", "NDVI", "2022_02_24_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_2 <- read_stars(here("imagery", "NDVI", "2022_02_28_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_3 <- read_stars(here("imagery", "NDVI", "2022_03_08_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_4 <- read_stars(here("imagery", "NDVI", "2022_03_16_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_5 <- read_stars(here("imagery", "NDVI", "2022_03_22_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_6 <- read_stars(here("imagery", "NDVI", "2022_04_05_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_7 <- read_stars(here("imagery", "NDVI", "2022_04_12_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_8 <- read_stars(here("imagery", "NDVI", "2022_04_20_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_9 <- read_stars(here("imagery", "NDVI", "2022_04_29_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_10 <- read_stars(here("imagery", "NDVI", "2022_05_03_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_11 <- read_stars(here("imagery", "NDVI", "2022_05_11_ndvi.tif")) %>% 
  setNames("ndvi")
#ndvi_12 <- read_stars(here("imagery", "NDVI", "2022_05_12_ndvi.tif")) %>% 
 # setNames("ndvi")
ndvi_13 <- read_stars(here("imagery", "NDVI", "2022_05_17_ndvi.tif")) %>% 
  setNames("ndvi")
ndvi_14 <- read_stars(here("imagery", "NDVI", "2022_05_29_ndvi.tif")) %>% 
  setNames("ndvi")


ndvi_all <- c(ndvi_1, ndvi_2, ndvi_3, ndvi_4, ndvi_5, ndvi_6, ndvi_7, ndvi_8, ndvi_9, ndvi_10, ndvi_11, ndvi_13, ndvi_14, along = 3) %>% 
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling VSSI:

```{r}
vssi_1 <- read_stars(here("imagery", "VSSI", "02_24_vssi.tif")) %>% 
  setNames("vssi")
vssi_2 <- read_stars(here("imagery", "VSSI", "02_28_vssi.tif")) %>% 
  setNames("vssi")
vssi_3 <- read_stars(here("imagery", "VSSI", "03_08_vssi.tif")) %>% 
  setNames("vssi")
vssi_4 <- read_stars(here("imagery", "VSSI", "03_16_vssi.tif")) %>% 
  setNames("vssi")
vssi_5 <- read_stars(here("imagery", "VSSI", "03_22_vssi.tif")) %>% 
  setNames("vssi")
vssi_6 <- read_stars(here("imagery", "VSSI", "04_05_vssi.tif")) %>% 
  setNames("vssi")
vssi_7 <- read_stars(here("imagery", "VSSI", "04_12_vssi.tif")) %>% 
  setNames("vssi")
vssi_8 <- read_stars(here("imagery", "VSSI", "04_20_vssi.tif")) %>% 
  setNames("vssi")
vssi_9 <- read_stars(here("imagery", "VSSI", "04_29_vssi.tif")) %>% 
  setNames("vssi")
vssi_10 <- read_stars(here("imagery", "VSSI", "05_03_vssi.tif")) %>% 
  setNames("vssi")
vssi_11 <- read_stars(here("imagery", "VSSI", "05_11_vssi.tif")) %>% 
  setNames("vssi")
#vssi_12 <- read_stars(here("imagery", "VSSI", "05_12_vssi.tif")) %>% 
 # setNames("vssi")
vssi_13 <- read_stars(here("imagery", "VSSI", "05_17_vssi.tif")) %>% 
  setNames("vssi")
vssi_14 <- read_stars(here("imagery", "VSSI", "05_29_vssi")) %>% 
  setNames("vssi")

vssi_all <- c(vssi_1, vssi_2, vssi_3, vssi_4, vssi_5, vssi_6, vssi_7, vssi_8, vssi_9, vssi_10, vssi_11, vssi_13, vssi_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling mARI:

```{r}
mari_1 <- read_stars(here("imagery", "mARI", "02_24_mari_avg.tif")) %>% 
  setNames("mari")
mari_2 <- read_stars(here("imagery", "mARI", "02_28_mari_avg.tif")) %>% 
  setNames("mari")
mari_3 <- read_stars(here("imagery", "mARI", "03_08_mari_avg.tif")) %>% 
  setNames("mari")
mari_4 <- read_stars(here("imagery", "mARI", "03_16_mari_avg.tif")) %>% 
  setNames("mari")
mari_5 <- read_stars(here("imagery", "mARI", "03_22_mari_avg.tif")) %>% 
  setNames("mari")
mari_6 <- read_stars(here("imagery", "mARI", "04_05_mari_avg.tif")) %>% 
  setNames("mari")
mari_7 <- read_stars(here("imagery", "mARI", "04_12_mari_avg.tif")) %>% 
  setNames("mari")
mari_8 <- read_stars(here("imagery", "mARI", "04_20_mari_avg.tif")) %>% 
  setNames("mari")
mari_9 <- read_stars(here("imagery", "mARI", "04_29_mari_avg.tif")) %>% 
  setNames("mari")
mari_10 <- read_stars(here("imagery", "mARI", "05_03_mari_avg.tif")) %>% 
  setNames("mari")
mari_11 <- read_stars(here("imagery", "mARI", "05_11_mari_avg.tif")) %>% 
  setNames("mari")
#mari_12 <- read_stars(here("imagery", "mARI", "05_12_mari_avg.tif")) %>% 
 # setNames("mari")
mari_13 <- read_stars(here("imagery", "mARI", "05_17_mari_avg.tif")) %>% 
  setNames("mari")
mari_14 <- read_stars(here("imagery", "mARI", "05_29_mari_avg.tif")) %>% 
  setNames("mari")

mari_all <- c(mari_1, mari_2, mari_3, mari_4, mari_5, mari_6, mari_7, mari_8, mari_9, mari_10, mari_11, mari_13, mari_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Opening and Compiling CRSI:

```{r}
crsi_1 <- read_stars(here("imagery", "CRSI", "02_24_crsi.tif")) %>% 
  setNames("crsi")
crsi_2 <- read_stars(here("imagery", "CRSI", "02_28_crsi.tif")) %>% 
  setNames("crsi")
crsi_3 <- read_stars(here("imagery", "CRSI", "03_08_crsi.tif")) %>% 
  setNames("crsi")
crsi_4 <- read_stars(here("imagery", "CRSI", "03_16_crsi.tif")) %>% 
  setNames("crsi")
crsi_5 <- read_stars(here("imagery", "CRSI", "03_22_crsi.tif")) %>% 
  setNames("crsi")
crsi_6 <- read_stars(here("imagery", "CRSI", "04_05_crsi.tif")) %>% 
  setNames("crsi")
crsi_7 <- read_stars(here("imagery", "CRSI", "04_12_crsi.tif")) %>% 
  setNames("crsi")
crsi_8 <- read_stars(here("imagery", "CRSI", "04_20_crsi.tif")) %>% 
  setNames("crsi")
crsi_9 <- read_stars(here("imagery", "CRSI", "04_29_crsi.tif")) %>% 
  setNames("crsi")
crsi_10 <- read_stars(here("imagery", "CRSI", "05_03_crsi.tif")) %>% 
  setNames("crsi")
crsi_11 <- read_stars(here("imagery", "CRSI", "05_11_crsi.tif")) %>% 
  setNames("crsi")
#crsi_12 <- read_stars(here("imagery", "CRSI", "05_12_crsi.tif")) %>% 
 # setNames("crsi")
crsi_13 <- read_stars(here("imagery", "CRSI", "05_17_crsi.tif")) %>% 
  setNames("crsi")
crsi_14 <- read_stars(here("imagery", "CRSI", "05_29_crsi.tif")) %>% 
  setNames("crsi")

crsi_all <- c(crsi_1, crsi_2, crsi_3, crsi_4, crsi_5, crsi_6, crsi_7, crsi_8, crsi_9, crsi_10, crsi_11, crsi_13, crsi_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

# Compiling NIR Bands

```{r}
nir_1 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_24.tif")) %>% 
  st_set_dimensions(3, values = seq(1:425)) %>% 
  filter(band %in% c(76:126)) %>% 
  split(3) %>% 
  setNames(c(76:126))

nir_2 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_28.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_3 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_08.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_4 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_16.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_5 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_22.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_6 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_05.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_7 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_12.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_8 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_20.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_9 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_10 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_03.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_11 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_11.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

#nir_12 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_12.tif")) %>%
#  st_set_dimensions(3, values = seq(1:425)) %>%
#  filter(band %in% c(76:126)) %>%
#  split(3) %>%
#  setNames(c(76:126))

nir_13 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_17.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_14 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))

nir_all <- c(nir_1, nir_2, nir_3, nir_4, nir_5, nir_6, nir_7, nir_8, nir_9, nir_10, nir_11, nir_13, nir_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date") %>% 
  st_warp(dest = crsi_all) 
```


# Principle Component Genertation

## Calculating PCs

```{r}
nir_pc_1 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_24.tif")) %>% 
  st_set_dimensions(3, values = seq(1:425)) %>% 
  filter(band %in% c(76:126)) %>% 
  split(3) %>% 
  setNames(c(76:126)) %>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_2 <- read_stars(here("imagery", "subset_images", "rfl_2022_02_28.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_3 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_08.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_4 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_16.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_5 <- read_stars(here("imagery", "subset_images", "rfl_2022_03_22.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_6 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_05.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_7 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_12.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_8 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_20.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_9 <- read_stars(here("imagery", "subset_images", "rfl_2022_04_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_10 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_03.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_11 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_11.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

#nir_12 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_12.tif")) %>%
#  st_set_dimensions(3, values = seq(1:425)) %>%
#  filter(band %in% c(76:126)) %>%
#  split(3) %>%
#  setNames(c(76:126))

nir_pc_13 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_17.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

nir_pc_14 <- read_stars(here("imagery", "subset_images", "rfl_2022_05_29.tif")) %>%
  st_set_dimensions(3, values = seq(1:425)) %>%
  filter(band %in% c(76:126)) %>%
  split(3) %>%
  setNames(c(76:126))%>% 
  as.data.frame() %>% 
  drop_na() %>% 
  prcomp()

```


## Applying PCs

```{r}
pc_1 <- predict(nir_1, nir_pc_1)

pc_2 <- predict(nir_2, nir_pc_2)

pc_3 <- predict(nir_3, nir_pc_3)

pc_4 <- predict(nir_4, nir_pc_4)

pc_5 <- predict(nir_5, nir_pc_5)

pc_6 <- predict(nir_6, nir_pc_6)

pc_7 <- predict(nir_7, nir_pc_7)

pc_8 <- predict(nir_8, nir_pc_8)

pc_9 <- predict(nir_9, nir_pc_9)

pc_10 <- predict(nir_10, nir_pc_10)

pc_11 <- predict(nir_11, nir_pc_11)

pc_13 <- predict(nir_13, nir_pc_13)

pc_14 <- predict(nir_14, nir_pc_14)
```

## Compling PCs

```{r}
merged_pc_1 <- merge(pc_1) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_2 <- merge(pc_2) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_3 <- merge(pc_3) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_4 <- merge(pc_4) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_5 <- merge(pc_5) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_6 <- merge(pc_6) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_7 <- merge(pc_7) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_8 <- merge(pc_8) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_9 <- merge(pc_9) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_10 <- merge(pc_10) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_11 <- merge(pc_11) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_13 <- merge(pc_13) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

merged_pc_14 <- merge(pc_14) %>% 
  st_as_stars() %>% 
  filter(attributes %in% c("PC3", "PC4", "PC5", "PC6", "PC7", "PC8", "PC9", "PC10", "PC11", "PC12", "PC13", "PC14")) %>% 
  split()

pc_all <- c(merged_pc_1, merged_pc_2, merged_pc_3, merged_pc_4, merged_pc_5, merged_pc_6, merged_pc_7, merged_pc_8, merged_pc_9, merged_pc_10, merged_pc_11, merged_pc_13, merged_pc_14, along = 3) %>%
  st_set_dimensions(3, values = lubridate::ymd(c("2022_02_24", "2022_02_28", "2022_03_08", "2022_03_16", "2022_03_22", "2022_04_05", "2022_04_12", "2022_04_20", "2022_04_29", "2022_05_03", "2022_05_11", "2022_05_17", "2022_05_29")), names = "date")
```

## Adjusting elevation data

```{r}
mllw_all <- mllw_all %>% 
  st_warp(dest = crsi_all) %>% 
  setNames("elevation")

```

# Extracting Values based on week

```{r}

soils_extract <- data.frame()

for (i in unique(lubridate::week(soils$paired_flight))){
  ndvi_week <- ndvi_all %>% 
    filter(lubridate::week(date) == i)
  
  mari_week <- mari_all %>% 
    filter(lubridate::week(date) == i)

  vssi_week <- vssi_all %>% 
    filter(lubridate::week(date) == i)
  
  crsi_week <- crsi_all %>% 
    filter(lubridate::week(date) == i)
  
  nir_week <- nir_all %>% 
    filter(lubridate::week(date) == i) %>% 
    st_as_sf()
  
  pc_week <- pc_all %>% 
    filter(lubridate::week(date) == i) %>% 
    st_as_sf()
  
  soil_week <- soils %>% 
    mutate(week = lubridate::week(paired_flight)) %>% 
    filter(week == i)
  
  ndvi_extract <- ndvi_week %>% 
    st_extract(soil_week)
  
  mari_extract <- mari_week %>% 
    st_extract(soil_week)
  
  vssi_extract <- vssi_week %>% 
    st_extract(soil_week)
  
  crsi_extract <- crsi_week %>% 
    st_extract(soil_week)
   
  nir_extract <- soil_week %>% 
    st_join(nir_week) %>% 
    select("76":"126") %>% 
    st_drop_geometry()
  
  pc_extract <- soil_week %>% 
    st_join(pc_week) %>% 
    select("PC3":"PC14") %>% 
    st_drop_geometry()
  
  soils_week <- bind_cols(soil_week, ndvi_extract$ndvi, mari_extract$mari, vssi_extract$vssi, crsi_extract$crsi, nir_extract, pc_extract) %>% 
    rename("ndvi" = "...12",
           "mari" = "...13",
           "vssi" = "...14",
           "crsi" = "...15")
    
  soils_extract <- rbind(soils_extract, soils_week)
}

#write.csv(soils_extract, file = here("data", "extracted_soils.csv"))
```

# Spectral and Environmental Variables

## Fitting the model to all variables

```{r}
set.seed(1234)

fit_all <- train(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi + crsi, 
                 data = soils_extract, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```


### Evaluating Model

```{r}
# using caret
fit_all_imp <- as.data.frame(fit_all$finalModel$importance)
fit_all_imp_scaled <- predict(preProcess(fit_all_imp, method = c("range")), fit_all_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = `%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = `%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(fit_all$finalModel$y, fit_all$finalModel$predicted)
```

### Ploting Model Fit

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", linewidth = 1.5)+
  geom_point(aes(y = fit_all$finalModel$y, x = fit_all$finalModel$predicted), color = "#E69512", size = 4)+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all$finalModel$y, x = fit_all$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 11)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 28, hjust = 0.5, face = "bold"),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#B4AA98"),
        axis.text = element_text(color = "#5b4f41", size = 18,),
        axis.title = element_text(color = "#5b4f41", size = 22, hjust = 0.5, face = "bold"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41", linewidth = 1))

#ggsave(filename = "all_plot_caret.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

all_corr <- cor(y= fit_all$finalModel$y, x = fit_all$finalModel$predicted)
all_corr

lm(fit_all$finalModel$y ~ fit_all$finalModel$predicted)
```

# Spectral Indices & PCs Only

## Fitting the model to spectral Indices

```{r}

fit_si <- train(electro_cond_mS_per_cm ~ ndvi + mari + vssi + crsi + `PC3` + `PC4` + `PC5` + `PC6` + `PC7` + `PC8` + `PC9` + `PC10` + `PC11` + `PC12` + `PC13` + `PC14`, 
                 data = soils_extract, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

### Evaluating Model

```{r}

fit_si_imp <- as.data.frame(fit_si$finalModel$importance)
fit_si_imp_scaled <- predict(preProcess(fit_si_imp, method = c("range")), fit_si_imp) %>% 
  rownames_to_column(var = "variable")


ggplot(fit_si_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_si_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_si_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
 theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "si_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(y = fit_si$finalModel$y, x= fit_si$finalModel$predicted)

```
 
### Observed v Predicted Plot

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y= fit_si$finalModel$y, x = fit_si$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_si$finalModel$y, x = fit_si$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 11)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Spectral Indices and PCs")+#plot title
theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "si_plot.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

# Elevation prediction
 
## Random Forest regression of Soil salinity and elevation
```{r}
fit_ele <- train(electro_cond_mS_per_cm ~ elevation, data = soils, method = "rf", trControl = trainControl(method = "cv", number = 10), importance = T)#random forest fitting of the data with stated equation
```

### Correlation

```{r}
elevation_cor <- cor(y = fit_ele$finalModel$y, x = fit_ele$finalModel$predicted)#calculate correlation coefficient
elevation_cor#call/print correlation value
```


### Observed v Predicted

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", linewidth = 1.5)+
  geom_point(aes(y = fit_ele$finalModel$y, x = fit_ele$finalModel$predicted), color = "#E69512", size = 3)+#data to make points from
  geom_smooth(method = "lm", aes(y = fit_ele$finalModel$y, x = fit_ele$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 11)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 28, hjust = 0.5, face = "bold"),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#B4AA98"),
        axis.text = element_text(color = "#5b4f41", size = 18,),
        axis.title = element_text(color = "#5b4f41", size = 22, hjust = 0.5, face = "bold"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41", linewidth = 1))

#ggsave(filename = "ele_plot.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```
 
# Compiling all raster data
 
```{r}
raster_all <- c(crsi_all, ndvi_all, vssi_all, mari_all, mllw_all)

``` 

# Predicting values

```{r}
all_pred <- predict(raster_all, fit_all, drop_dimensions = F)

```

# Making a Map

```{r}
all_pred_feb <- all_pred %>% 
  filter(date == "2022-02-24") %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE)

ggplot()+
  geom_stars(data = all_pred_feb, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "desert", n = 2, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Northing",#labels
       x = "Easting",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```

# Interactive Map

```{r}
mapview::mapview(all_pred_feb)
```



# Predicting values

```{r}
fit_si <- train(electro_cond_mS_per_cm ~ ndvi + mari + vssi + crsi, 
                 data = soils_extract, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

si_pred <- predict(raster_all, fit_si, drop_dimensions = F)

```

# Making a Map

```{r}
si_pred_feb <- si_pred %>% 
  filter(date == "2022-05-17") %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE)


ggplot()+
  geom_stars(data = si_pred_feb, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "desert", n = 2, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Northing",#labels
       x = "Easting",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("May 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```

# Predicting values

```{r}
ele_pred <- predict(raster_all, fit_ele, drop_dimensions = F)

```

# Making a Map

```{r}
ele_pred_feb <- ele_pred %>% 
  filter(date == "2022-05-17") %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE)

ggplot()+
  geom_stars(data = ele_pred_feb, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "desert", n = 2, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Northing",#labels
       x = "Easting",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("May 17, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```



















###################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################








# TESTING COLLINEARITY AND ALTERATIVE MODEL FITS

## Collinearity

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi + crsi, data = soils_extract)

car::vif(model)

```

Based on the VIF collinearity data, CRSI is highly correlated with the other variables and should potentially be removed from the model, NDVI and mARI seem to have moderate collinearity as well, potential removal of ndvi may be a possible solution. 

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi, data = soils_extract)

car::vif(model)
```

With CRSI removed, VIF values for mARI and NDVI are reduced, but they seem to be the two collinear values. Might be worth testing model fits between a model excluding CRSI and NDVI and just excluding CRSI.

```{r}
model <- lm(electro_cond_mS_per_cm ~ elevation + mari + vssi, data = soils_extract)

car::vif(model)
```

## Testing Exclusion of Spectral Indices (due to collinearity)

```{r}
set.seed(1234)

fit_all_nc <- train(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi, 
                 data = soils_extract, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```


### Evaluating Model

```{r}
# using caret
fit_all_nc_imp <- as.data.frame(fit_all_nc$finalModel$importance)
fit_all_nc_imp_scaled <- predict(preProcess(fit_all_nc_imp, method = c("range")), fit_all_nc_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_all_nc_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_all_nc_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_all_nc_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))
  
#ggsave(filename = "all_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(fit_all_nc$finalModel$y, fit_all_nc$finalModel$predicted)
```

### Ploting Model Fit

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted", linewidth = 1.5)+
  geom_point(aes(y = fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted), color = "#E69512", size = 4)+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 11)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 28, hjust = 0.5, face = "bold"),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#B4AA98"),
        axis.text = element_text(color = "#5b4f41", size = 18,),
        axis.title = element_text(color = "#5b4f41", size = 22, hjust = 0.5, face = "bold"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41", linewidth = 1))

#ggsave(filename = "all_plot_nc.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

all_corr <- cor(y= fit_all_nc$finalModel$y, x = fit_all_nc$finalModel$predicted)
all_corr

lm(fit_all_nc$finalModel$y ~ fit_all_nc$finalModel$predicted)
```



```{r}
all_pred_nc <- predict(raster_all, fit_all_nc, drop_dimensions = F, na.action = na.omit)
```


```{r}
all_pred_crop <- all_pred_nc  %>% 
  st_crop(y = boundary, crop = TRUE, as_points = FALSE) %>% 
  filter(date == "2022-02-24")

ggplot()+
  geom_stars(data = all_pred_crop, aes(x = x, y = y, fill = prediction), na.action = na.omit) +
  scale_fill_gradientn(colors = rev(cal_palette(name = "desert", n = 10, type = "continuous")))+
  coord_sf(crs = 32116)+
  labs(y = "Northing",#labels
       x = "Easting",
       fill = "Electrical 
Conductivity 
(mS/cm)")+
  ggtitle("February 24, 2022")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 16, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41"),
        axis.title = element_text(color = "#5b4f41"),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"),
        legend.title = element_text(color = "#5b4f41"),
        legend.text = element_text(color = "#5b4f41"))
```

```{r}
mapview(all_pred_crop)
```


########################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################



## Including PCs

```{r}

fit_pc <- train(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi + crsi + `PC3` + `PC4` + `PC5` + `PC6` + `PC7` + `PC8` + `PC9` + `PC10` + `PC11` + `PC12` + `PC13` + `PC14`, 
                 data = soils_extract, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)

```

### Evaluating Model

```{r}
# using caret
fit_pc_imp <- as.data.frame(fit_pc$finalModel$importance)
fit_pc_imp_scaled <- predict(preProcess(fit_pc_imp, method = c("range")), fit_pc_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_pc_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_pc_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_pc_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_si_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

cor(fit_pc$finalModel$y, fit_pc$finalModel$predicted)
```

### Ploting Model Fit

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_pc$finalModel$y, x = fit_pc$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_pc$finalModel$y, x = fit_pc$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 11)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation, Indices, and PC")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "all_si_plot.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot

```






####################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################################









# Comparing which variables are most important by landcover type

## Bare soil

```{r}

soils_soils <- soils_extract %>% 
  filter(landcover == "soil")


fit_soil <- train(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi + crsi, 
                 data = soils_soils, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)


```


### Evaluating Model

```{r}
# using caret
fit_soil_imp <- as.data.frame(fit_soil$finalModel$importance)
fit_soil_imp_scaled <- predict(preProcess(fit_soil_imp, method = c("range")), fit_soil_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_soil_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_soil_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_soil_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "bare_soil_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

### Ploting Model Fit

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_soil$finalModel$y, x = fit_soil$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_soil$finalModel$y, x = fit_soil$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(3, 14)) +
  scale_y_continuous(expand = c(0, 0), limits = c(3, 14))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

cor(fit_soil$finalModel$y, fit_soil$finalModel$predicted)

#ggsave(filename = "bare_soil_plot.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```


## Vegetated

```{r}

soils_veg <- soils_extract %>% 
  filter(landcover == "vegetated")


fit_veg <- train(electro_cond_mS_per_cm ~ elevation + ndvi + mari + vssi + crsi, 
                 data = soils_veg, 
                 method = "rf", 
                 trControl = trainControl(method = "cv", number = 10), 
                 importance = TRUE)


```

### Evaluating Model

```{r}
# using caret
fit_veg_imp <- as.data.frame(fit_veg$finalModel$importance)
fit_veg_imp_scaled <- predict(preProcess(fit_veg_imp, method = c("range")), fit_veg_imp) %>% 
  rownames_to_column(var = "variable")

ggplot(fit_veg_imp_scaled) +
  geom_point(aes(x = IncNodePurity, y = variable), color = "#E69512")+
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = IncNodePurity), color = "#E69512", size = 1.2)+
  geom_point(aes(x = fit_veg_imp_scaled$`%IncMSE`, y = variable), color ="#3A5D3D")+ 
  geom_segment(aes(y= variable, x = 0, yend = variable, xend = fit_veg_imp_scaled$`%IncMSE`), color = "#3A5D3D", linetype = "dashed")+
   labs(y = "Variables",#labels
       x = "Importance (Scaled)")+
  ggtitle("Random Forest Variable Importance")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

#ggsave(filename = "veg_imp.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

### Ploting Model Fit

```{r}
ggplot()+#ggplot object
  geom_abline(slope = 1, intercept = 0, linetype = "dotted")+
  geom_point(aes(y = fit_veg$finalModel$y, x = fit_veg$finalModel$predicted), color = "#E69512")+#data to make points from
  geom_smooth(method = "lm", aes(y= fit_veg$finalModel$y, x = fit_veg$finalModel$predicted), color =  "#3A5D3D")+#smooth line data
  scale_x_continuous(expand = c(0, 0), limits = c(0, 12)) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 12))+
  labs(y = "Actual Electrical Conductivity (mS/cm)",#labels
       x = "Predicted Electrical conductivity (mS/cm)")+
  ggtitle("Elevation and Indices")+#plot title
  theme(plot.title = element_text(color = "#5b4f41", size = 24, hjust = 0.5),#plot themes
        plot.background = element_rect("white"),
        panel.background = element_rect("#faf7f2"),
        panel.grid = element_line(linetype= "longdash", color = "#f0ece1"),
        axis.text = element_text(color = "#5b4f41", size = 16),
        axis.title = element_text(color = "#5b4f41", size = 20, hjust = 0.5),
        strip.background = element_rect("#f8f8f8"),
        strip.text = element_text(color = "#5b4f41"),
        axis.line = element_line(color = "#5b4f41"))

cor(fit_veg$finalModel$y, fit_veg$finalModel$predicted)

#ggsave(filename = "veg_plot.png", plot = last_plot(), width = 12, height = 8, device = "png", dpi = 500) #write an image of the plot
```

